---
title: "p8105_hw3_yf2605"
author: "Yi_Fang"
date: '`r Sys.Date()`'
output: github_document
---
# Setting

libraries

```{r}
library(tidyverse)
library(p8105.datasets)
```

picture display setting

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

# Problem 1

## load Instacart data

```{r}
data("instacart")
```

## Description:

The Instacart dataset contains `r ncol(instacart)` variables and `r nrow(instacart)` records.

The variables include:
`r colnames(instacart)`

## Questions:

*   How many aisles are there, and which aisles are the most items ordered from?

```{r}
n_aisles = 
  instacart %>% 
  distinct(aisle) %>% 
  nrow()

most_item_aisle = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_items = n()) %>% 
  slice(which.max(n_items))
```

`r n_aisles` aisles are there, and `r most_item_aisle$aisle` aisle is where the most items ordered from.

*   Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_items = n()) %>% 
  filter(n_items > 10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_items)) %>% 
  ggplot(aes(aisle, y = n_items)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

*   Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}

instacart %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_items = n()) %>% 
  filter(aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>% 
  arrange(desc(n_items)) %>% 
  slice(1:3) %>%
  knitr::kable()
```

*   Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart %>% 
  group_by(order_dow, product_name) %>% 
  summarize(mean_hour = mean(order_hour_of_day, na.rm=TRUE)) %>% 
  filter(product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable(digits = 2)
```

# Problem 2

## read and tidy the accel data set

* read the data set
* clean variable names
* convert variables: activity_1-activity_1440 into longer version
* add factor variable: weekend, 1 for weekend, 0 for weekday
* convert variable: day as factor variable
* convert activity counts as integer

```{r}
accel = 
  read_csv(file = 'data/accel_data.csv') %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = 'minute_from_midnight',
    names_prefix = 'activity_',
    values_to = 'activity_counts'
  ) %>% 
  mutate(
    weekend = as.factor(if_else(day %in% c('Saturday', 'Sunday'), 1, 0)),
    day = as.factor(day),
    activity_counts = as.integer(activity_counts)
    )
```

## Description

The accel dataset contains `r ncol(accel)` variables and `r nrow(accel)` records.

The variables include:
`r colnames(accel)`

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

# Problem 3

This problem uses the NY NOAA data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package using:

```{r}
data("ny_noaa")
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
