---
title: "p8105_hw3_yf2605"
author: "Yi_Fang"
date: '`r Sys.Date()`'
output: github_document
---
# Setting

libraries

```{r}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)
```

# Problem 1

## load Instacart data

```{r}
data("instacart")
```

## Description:

The Instacart dataset contains `r ncol(instacart)` variables and `r nrow(instacart)` records.

The variables include:

* `r colnames(instacart)`

code:

* `order_id`: order identifier
* `product_id`: product identifier
* `add_to_cart_order`: order in which each product was added to cart
* `reordered`: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
* `user_id`: customer identifier
* `eval_set`: which evaluation set this order belongs in (Note that the data for use in this class is exclusively from * the “train” eval_set)
* `order_number`: the order sequence number for this user (1=first, n=nth)
* `order_dow`: the day of the week on which the order was placed
* `order_hour_of_day`: the hour of the day on which the order was placed
* `days_since_prior_order`: days since the last order, capped at 30, NA if order_number=1
* `product_name`: name of the product
* `aisle_id`: aisle identifier
* `department_id`: department identifier
* `aisle`: the name of the aisle
* `department`: the name of the department

summary:

```{r}
summary(instacart) %>% 
  knitr::kable()
```


## Questions:

### How many aisles are there, and which aisles are the most items ordered from?

```{r}
n_aisles = 
  instacart %>% 
  distinct(aisle) %>% 
  nrow()

most_item_aisle = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_items = n()) %>% 
  arrange(desc(n_items)) %>% 
  slice(1:4)
```

`r n_aisles` aisles are there, and `r most_item_aisle$aisle` aisles are where the most items ordered from.

### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_items = n()) %>% 
  filter(n_items > 10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_items)) %>% 
  ggplot(aes(x = aisle, y = n_items)) +
  geom_col() +
  labs(
    title = "Number of items ordered in each aisle",
    x = "Aisle",
    y = "Number of items ordered",
    caption = "Only show aisles with more than 10000 items ordered"
  )+
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
    )
```

### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}

instacart %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_items = n()) %>% 
  filter(aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>% 
  arrange(desc(n_items)) %>% 
  slice(1:3) %>%
  knitr::kable()
```

### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart %>% 
  group_by(order_dow, product_name) %>% 
  summarize(mean_hour = mean(order_hour_of_day, na.rm=TRUE)) %>% 
  filter(product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable(digits = 2)
```

# Problem 2

## read and tidy the accel data set

* read the data set
* clean variable names
* convert variables: activity_1-activity_1440 into longer version
* add factor variable: weekend, 1 for weekend, 0 for weekday
* convert variable: day as factor variable
* convert variables: minute_from_midnight as integer

```{r}
accel = 
  read_csv(file = 'data/accel_data.csv') %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = 'minute_from_midnight',
    names_prefix = 'activity_',
    values_to = 'activity_counts'
  ) %>% 
  mutate(
    weekend = as.factor(if_else(day %in% c('Saturday', 'Sunday'), 1, 0)),
    day = factor(day, levels=c('Sunday', 'Monday', 'Tuesday', 'Wednesday', 
                               'Thursday', 'Friday', 'Saturday')),
    minute_from_midnight = as.integer(minute_from_midnight),
    )
```

## Description

The accel data set contains `r ncol(accel)` variables and `r nrow(accel)` records.

The variables include:

* `r colnames(accel)`

summary:

```{r}
summary(accel) %>% 
  knitr::kable()
```

## Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

* calculate total activity counts for each day
* make a table

```{r}
total_activity = 
  accel %>% 
  group_by(week, day) %>% 
  summarize(
    total_activity = sum(activity_counts)
  )

total_activity %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activity
  ) %>% 
  knitr::kable(digits = 2)
```

Trends:

* Activity counts on Mondays increases from week 1 to week 3 and decreases from week 3 to week 5.

## Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel %>% 
  ggplot(aes(x = minute_from_midnight, y = activity_counts, color = day)) +
  geom_smooth(se = FALSE) +
  scale_x_continuous(
    breaks = seq(0, 1440, by = 60),
    labels = seq(0, 24)
    ) +
  labs(
    title = "24-hour activity time courses for each day of the week",
    x = "Hour",
    y = "Activity counts",
    color='Day of the week'
  )+
  theme(
    legend.position = "bottom", 
    plot.title = element_text(hjust = 0.5)
    )
```

conclusions:

* Most activities are carried out between 8:00 to 23:00. 

* There is one peak of activity on Sundays and one on Fridays. 

* On Sundays the activity peak is between 10:00-11:00 and on Fridays it is between 20:00-22:00.

# Problem 2

## Load data

```{r}
data("ny_noaa")
```

## Description

The ny_noaa data set contains `r ncol(ny_noaa)` variables and `r nrow(ny_noaa)` records.

The variables include:

* `r colnames(ny_noaa)`

code:

* `id`: Weather station ID
* `date`: Date of observation
* `prcp`: Precipitation (tenths of mm)
* `snow`: Snowfall (mm)
* `snwd`: Snow depth (mm)
* `tmax`: Maximum temperature (tenths of degrees C)
* `tmin`: Minimum temperature (tenths of degrees C)

summary:

```{r}
summary(ny_noaa) %>% 
  knitr::kable()
```

Missing data:

* `r round(100 * sum(is.na(ny_noaa$prcp))/nrow(ny_noaa),2)` % of prcp is missing.
* `r round(100 * sum(is.na(ny_noaa$snow))/nrow(ny_noaa),2)` % of snow is missing.
* `r round(100 * sum(is.na(ny_noaa$snwd))/nrow(ny_noaa),2)` % of snwd is missing.
* `r round(100 * sum(is.na(ny_noaa$tmax))/nrow(ny_noaa),2)` % of tmax is missing.
* `r round(100 * sum(is.na(ny_noaa$tmin))/nrow(ny_noaa),2)` % of tmin is missing.


## Questions

### Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

* Create separate variables for year, month, and day.
* convert numeric month to month name
* convert tmin and tmax to numeric variable and set unit to degrees Celsius.
* set prcp unit to mm.

```{r}
ny_noaa_cleaned = 
  ny_noaa %>% 
  separate(date, c('year', 'month', 'day'), convert = TRUE) %>% 
  mutate(
    month = month.name[month], 
    tmax = as.numeric(tmax) / 10,
    tmin = as.numeric(tmin) / 10,
    prcp = prcp / 10
  )
```


```{r}
most_obs_snow =
  ny_noaa_cleaned %>% 
  group_by(snow) %>% 
  summarise(n_obs = n()) %>% 
  drop_na(snow) %>% 
  arrange(desc(n_obs)) %>% 
  slice(1:3)
```

For snowfall, the most commonly observed values are `r most_obs_snow$snow` mm.

### Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
ny_noaa_cleaned %>% 
  filter(month %in% c('January', 'July')) %>% 
  group_by(year, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm=TRUE)
  ) %>% 
  ggplot(aes(x = year, y = mean_tmax)) +
  geom_point() +
  geom_line() +
  facet_grid(~month) +
  scale_y_continuous(breaks = seq(-7, 30, by = 2)) +
  labs(
    title = "Average max temperature in January and in July from 1981 to 2010",
    x = "Year",
    y = "Average max temperature in Celsius",
  )+
  theme(plot.title = element_text(hjust = 0.5))
```

* Average max temperature in January is lower than that in July.
* The average max temperature in January is fluctuating between -5 to 5 degrees Celsius.
* The average max temperature in January is fluctuating between 24 to 29 degrees Celsius.
* No outliers.

### Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
tmax_tmin_plot = 
  ny_noaa_cleaned %>% 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_smooth() +
  labs(
    title = "Maximum vs Minimum Daily temperature",
    x = "Daily Maximum temperature(°C)",
    y = "Daily Minimum temperature(°C)",
  ) +
  theme(plot.title = element_text(hjust = 0.5))

snow_plot = 
  ny_noaa_cleaned %>% 
  filter(snow > 0, snow < 100) %>%
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(x = snow, y = year)) +
  geom_density_ridges(scale = .90) +
  labs(
    title = "Distribution of snowfall by year",
    x = "Snowfall (mm)",
    y = "Year",
    caption = 'Only shows snowfall values greater than 0 and less than 100'
  )+
  theme(plot.title = element_text(hjust = 0.5))

tmax_tmin_plot + snow_plot
```

